### 2019-12-11 rnni_convexity.tex
rm commeted out tex with some ideas and importantly LATTICES -- @lenacoll please check you don't need these and add the lattices section back

\documentclass{amsart}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage[misc]{ifsym}
\usepackage[foot]{amsaddr}
\usepackage[hidelinks]{hyperref}
\usepackage{csquotes}
\usepackage[notref,notcite]{showkeys}
\usepackage[style=authoryear,ibidtracker=false,uniquename=false,giveninits=true,terseinits=true,maxbibnames=5,backend=biber]{biblatex}
\renewbibmacro{in:}{}
\addbibresource{rnni_convexity.bib}

\newcommand{\np}{\mathcal{NP}}
\newcommand{\parent}{\mathrm{parent}}
\newcommand{\mrca}{\mathrm{mrca}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\nni}{\mathrm{NNI}}
\newcommand{\rnni}{\mathrm{RNNI}}
\newcommand{\rnniu}{\mathrm{RNNIu}}
\newcommand{\tbr}{\mathrm{TBR}}
\newcommand{\spr}{\mathrm{SPR}}
\newcommand{\csort}{\textsc{Caterpillar Sort}}
\newcommand{\findpath}{\textsc{FindPath}}
\newcommand{\mdtree}{\textsc{MDTree}}

\newtheorem{definition}{Definition}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{conjecture}[definition]{Conjecture}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}

\graphicspath{{figures/}}

\sloppy

\title[Ranked Nearest Neighbour Intarchange]{Geometry of Ranked Nearest Neighbour Interchange Space of Phylogenetic Trees}
\date{\today}
\author{Lena Collienne\textsuperscript{1}}
\email{lena.collienne@postgrad.otago.ac.nz}
\address{\textsuperscript{1}Department of Computer Science, University of Otago, New Zealand}
\author{Kieran Elmes\textsuperscript{1}}
\email{kelmes@cs.otago.ac.nz}
\author{Mareike Fischer\textsuperscript{2}}
\email{email@mareikefischer.de}
\address{\textsuperscript{2}Institute for Mathematics and Inofrmatics, University of Greifswald, Germany}
\author{David Bryant\textsuperscript{3}}
\email{david.bryant@otago.ac.nz}
\address{\textsuperscript{3}Department of Mathematics and Statistics, University of Otago, New Zealand}
\author{Alex Gavryushkin\textsuperscript{1, \Letter}}
\email{\textsuperscript{\Letter}alex@biods.org}

\begin{document}

\begin{abstract}
In this paper we study the graph of ranked phylogenetic trees where the adjacency relation is given by a local rearrangement of the tree structure.
Our work is motivated by tree inference algorithms, such as maximum likelihood and Markov Chain Monte Carlo methods, where the geometry of the search space plays a central role for efficiency and practicality of the optimisation.
We hence focus on understanding the geometry of the space (graph) of ranked trees, the so-called ranked nearest neighbour interchange ($\rnni$) graph.
For example, we find exact estimates of the radius and diameter of the space, improving the best previously known estimates.
Since the $\rnni$ graph is a generalisation of the classical nearest neighbour interchange ($\nni$) graph to ranked phylogenetic trees, we compare geometric and algorithmic properties of the two graphs.
Surprisingly, we discover that both geometric and algorithmic properties of $\rnni$ and $\nni$ are quite different.
For example, we establish convexity of certain natural subspaces in $\rnni$, which are not convex is $\nni$.
Furthermore, our results suggest that the complexity of computing distances in the two graphs is different.
\end{abstract}

\maketitle

The nearest neighbour interchange ($\nni$) graph, defined on the set of phylogenetic trees with adjacency relation given by the interchange operation of two sister clades, has been known in mathematical biology literature for nearly 50 years \autocite{Robinson1971-ql,Moore1973-kk}.
Considered with the metric given by the length of a shortest path (graph-distance), this graph becomes a metric space.
Its geometry has been extensively studied \autocite{Dasgupta2000-xa, Li1996-zw, Gordon2013-fw, De_Jong2016-al}.
An important property of the $\nni$ graph is that computing distances is NP-hard \autocite{Dasgupta2000-xa}.
As a result no algorithm exists to compute the distance in practical time.
A consequence is that tree search and sampling algorithms pose a significant challenge even for moderately sized trees
\autocite{Whidden2016-kl}.

More recent advances in computational phylogenetics introduce various classes of molecular clock models \autocite{Yoder2000-ks,Drummond2006-nl,Drummond2010-yf} and made computational inference of phylogenetic time-trees possible \autocite{Ronquist2003-eq, Bouckaert2018-yr, Hadfield2018-xp}.
However, mathematical challenges that come with this seemingly inessential change in parametrisation (genomic distance VS time distance) of trees have only recently been brought to attention \autocite{Gavryushkin2016-uu}.
These differences motivated \textcite{Gavryushkin2018-ol} to propose an extension of the $\nni$ graph to the class of discrete time-trees.
The simplest such extension introduces the $\rnni$ graph on the set of ranked phylogenetic trees.
Considered with the graph-distance $\rnni$ becomes a metric space and inherits the geometric and algorithmic challenges that the $\nni$ space has been traditionally facing.
Surprisingly, most of them cannot be settled by directly translating results or applying techniques developed for $\nni$ \autocite{Gavryushkin2018-ol}.

In this paper, we consider the $\rnni$ space on ranked phylogenetic trees with all taxa being of equal rank.
An example of such a tree is depicted in Figure~\ref{fig:ranked_tree}.
In the terminology of \autocite{Gavryushkin2018-ol}, the space considered in this paper is the space of ranked ultrametric phylogenetic trees $\rnniu$.
We postpone accurate definitions until later in this section.

\begin{figure}[H]
\centering
\includegraphics[width=0.25\textwidth]{ranked_tree}
\vspace{12pt}
\caption{A ranked tree with $6$ taxa.}
\label{fig:ranked_tree}
\end{figure}

In line with the research programme proposed in \autocite{Gavryushkin2018-ol}, we investigate the geometry and algorithmic complexity of the $\rnni$ space.
Specifically, in this paper we establish the exact radius and diameter of the space (Section~\ref{section:diameter}).
We also show that the subset of caterpillar trees is convex in $\rnni$ (Section~\ref{section:caterpillar_convex}), thus settling one of the open problems proposed in \autocite{Gavryushkin2018-ol}.
For establishing these geometric properties we are using algorithms that will be introduced in Section~\ref{section:algorithms}.
We will in particular provide an approximate algorithm that computes exact distances for small trees.
The question of whether there exists a polynomial algorithm for computing the $\rnni$ distance remains an open problem.

In the rest of this chapter we formally introduce the terminology used in this paper.

A \emph{ranked phylogenetic tree} is a pair consisting of a rooted binary phylogenetic tree $T$ on the set $X = \{1, \ldots, n\}$ of \emph{taxa} for $n \in \mathbb N$, and a (total) rank function that maps all leaves of $T$ to $0$, all internal nodes of $T$ onto elements of the set $\{1, \ldots, n-1\}$, and respects the partial order on the nodes given by the tree.
The latter means that if $u$ and $v$ are two internal nodes of $T$ such that there exists a path from a taxon $x \in X$ to the root which first passes through $u$ and then through $v$ then $\rank(u) < \rank(v)$.
Ranked trees $(T_1, \rank_1)$ and $(T_2, \rank_2)$ are different if trees $T_1$ and $T_2$ are different or $T_1 = T_2$ and $\rank_1 \neq \rank_2$.
Since all trees in this paper are ranked trees, we will abuse the notation and drop the rank function from the notation.
We will also simply say trees to mean ranked phylogenetic trees.
For a tree $T$, we will use $\rank_T$ to refer to its ranking.
A ranked tree $T$ without its leaf labels will be called \emph{tree topology}.

Our definition of a ranked tree implies a natural notion of edge length -- we call $|\rank_T(v) - \rank_T(u)|$ the length of an edge $(u, v)$.

Now we are ready to introduce the tree space which is the subject of study in this paper, the $\rnni$ graph.

The vertex set of the $\rnni$ graph is the set of all ranked trees on $n$ taxa.
We introduce two types of operation ($\rnni$ \emph{moves}) on trees (see Figure~\ref{fig:RNNI}) and say that two trees are adjacent in the $\rnni$ graph if they are connected by an operation of either type.
The first type of operation is called a \emph{rank move} and defined by swapping the ranks of two internal nodes that are not adjacent in the tree.
Formally, if $u$ and $v$ are nodes of a tree $T$ such that $|\rank_T(u) - \rank_T(v)| = 1$ and $(u, v)$ is not an edge in $T$ then the tree $R$ obtained from $T$ by only changing $\rank_R(u) = \rank_T(v)$ and $\rank_R(v) = \rank_T(u)$ is said to be obtained by a rank move.
The second type of operation is called an $\nni$ \emph{move} and defined in the usual way, that is, two trees $T$ and $R$ are said to be connected by an $\nni$ move if there exist internal edges $e$ in $T$ and $f$ in $R$ both of length one such that the trees obtained by shrinking $e$ and $f$ to internal nodes coincide.

We use the notation $d(T, R)$ throughout the paper to denote the graph distance between trees in $\rnni$, that is, $d(T, R)$ is the length of a shortest $\rnni$ path between tree $T$ and $R$.

\begin{figure}[H]
\centering
\includegraphics[width=0.8\textwidth]{RNNI}
\vspace{12pt}
\caption{Two types of operation that define edges of the $\rnni$ graph.
The rank move swaps the rank of the two highlighted nodes and the $\nni$ moves are performed on the blue edges.}
\label{fig:RNNI}
\end{figure}

The rest of this paper is organised as follows.
In Section~\ref{section:algorithms} we introduce three algorithms for exploring the $\rnni$ graph.
We use these algorithms in Section~\ref{section:geometry} to establish several geometrical properties of this graph, such as its diameter and radius.
Furthermore, we will establish in Section~\ref{section:caterpillar_convex} that a certain subspace of $\rnni$ is convex and design a polynomial-time algorithm for computing $\rnni$-distances in that subspace.
This result suggests that the complexity of computing distances in $\rnni$ is different from that in the classical $\nni$ space.
We will discuss this further in Section~\ref{section:cluster_theorem}, where we disprove a conjecture of \textcite{Gavryushkin2018-ol} and suggest its weaker version a plausible alternative.
We will finish the paper with a discussion and open problems.

\section{Algorithms}
\label{section:algorithms}

In this section we introduce three algorithms, $\findpath$ (Section~\ref{section:alg_findpath}), $\csort$ (Section~\ref{section:alg_csort}), and $\mdtree$ (Section~\ref{section:alg_mdtree}).
$\findpath$ computes a $\rnni$ path between trees.
$\csort$ computes a shortest $\rnni$ path between caterpillar trees (also known as ladder trees, see later in this section for accurate definitions).
$\mdtree$ maximises the dissimilarity to find a remote tree given an arbitrary $\rnni$ tree.
While these algorithms are interesting on their own (e.g.\ in simulation studies), they will also form an important ingredient to study the geometry of the $\rnni$ space later in this paper.
For example, $\mdtree$ will be an important tool for finding the radius of the $\rnni$ graph, as the algorithms computes a tree with maximum distance if the input tree is a caterpillar tree.
We will prove this fact in Section~\ref{section:diameter}.

The algorithm $\findpath$ can be used to approximate the $\rnni$ distance.
We will study the accuracy of this approximation by computing the exact $\rnni$ distance for small trees.
Specifically, we will use the algorithm for computing the complete $\rnni$ graph that is given in \autocite[their Section 3.3]{Gavryushkin2018-ol}.
After pre-computing the graph, we use Seidel's algorithm~\autocite{seidel_all-pairs-shortest-path_1995} for computing distances.
However, the large number of vertices in $\rnni$ is an obstacle when it comes to computing distances for larger trees.
Since there are $\frac{n!(n-1)!}{2^{n-1}}$ vertices in the $\rnni$ graph \autocite{Gavryushkin2018-ol}, we compute the $\rnni$ graph on up to seven taxa for our analyses.
This computational result will also be used later in the paper to support Conjecture~\ref{conjecture:cluster_theorem}, a modification of Conjecture~9 in \autocite{Gavryushkin2018-ol}, which we prove to be false in this paper.

%TODO: Make a zenodo dio for the repository before final publication
We have implemented these algorithms in a combination of Python and C, and the source code is available at \autocite{Collienne2019}.

\subsection{$\findpath$}
\label{section:alg_findpath}

In this section we present $\findpath$, an efficient algorithm for computing paths, and therefore approximating distances, between trees in $\rnni$.

Before introducing the algorithm we need the following definitions.
Each node $v$ of a tree defines a \emph{cluster} $C$, which is the set of taxa descending from $v$.
We then say that $v$ \emph{induces} $C$.
\emph{Trivial clusters} are clusters induced by leaves (contain just one element each) and the root of the tree (contains all taxa).
As all trees on $n$ taxa share all trivial clusters, we will use the notion cluster to mean non-trivial clusters.
Note that the list of all non-trivial clusters of a tree, sorted according to the rank of the corresponding node, defines the tree unambiguously.
We will refer to this list of $n-2$ clusters for a given tree on $n$ taxa as a \emph{cluster representation} of the tree.

Let $T$ and $R$ be arbitrary trees and $[C_1, \ldots, C_{n-2}]$ be the cluster representation of $R$ (the ``destination'' tree).
Then $C_i$ is induced by the node of rank $i$ for $i = 1, \ldots, n-2$.
$\findpath$ computes a path by iteratively extending a sequence $p$ of trees starting from $T$.
The algorithm terminates when $p$ is an $\rnni$ path from $T$ to $R$.

\todo{Alex: go through the following two paragraphs and proof of L1}
In step $k$ of $\findpath$ we consider the cluster $C_k$ for $k = 1, \ldots, n-2$.
$p$ is extended in step $k$ by considering the last tree $T'$ on $p$ and finding the node in $T'$ that induces the smallest cluster containing all elements of $C_k$.
This node is denoted by $\mrca_{T'}(C_k)$ (short for \emph{most recent common ancestor}).
By performing an $\rnni$ move that decreases the rank of $\mrca_{T'}(C_k)$, we receive a tree that is then appended to $p$.
Now we extend $p$ by repeatedly performing $\rnni$ moves on the last tree of $p$ that move the most recent common ancestor of $C_k$ down.
In the end of step $k$, $C_k$ is induced by the internal node of rank $k$ in the last tree of $p$.

By following this procedure, the last of $p$ after step $n-2$ will be $R$, as it contains all clusters of $R$.
It is important but easy to see that the output path of $\findpath$ is unique for a given pair of trees, which means that it is a deterministic algorithm.
This is due to the fact that there exists exactly one $\rnni$ move that decreases the rank of the most recent common ancestor of a subset $C_k$ of the taxon set of a tree as long as $C_k$ is not induced by an internal node of that tree.

\begin{algorithm}[H]
\caption{$\findpath$($T,R$)}
\label{alg:find_path}
\begin{algorithmic}[1]
\STATE $T' := T$, $p := [T']$
\FOR {$k = 1, \dots, n-2$}
  \STATE Let $C_k$ be the cluster induced by the node with rank $k$ in $R$ \label{alg:find_path:line:cluster}
  \WHILE {$\rank_{T'}(\mrca_{T'}(C_k))>k$}
    \STATE Update $T'$: Decrease $\rank_{T'}(\mrca_{T'}(C_k))$ by an $\rnni$ move \label{alg:findpath:line:move_set_down}
    \STATE $p = p+T'$
  \ENDWHILE
\ENDFOR
\RETURN $p$
\end{algorithmic}
\end{algorithm}

The following lemma shows that $\findpath$ is correct, that is, the algorithm returns a path between the two input trees.
To prove this we have to establish that there always exists a unique $\rnni$ move required in Line~\ref{alg:findpath:line:move_set_down} of the algorithm.

\begin{lemma}
Let $T = [C_1, \ldots, C_{k-1}, B_{k}, \ldots, B_{n-2}]$ and $R = [C_1, \ldots, C_{k-1}, C_k, \ldots, C_{n-2}]$ be two trees where $C_k \neq B_k, \ldots, B_{n-2}$.
Then there exists exactly one $\rnni$ move on $T$ that results in a tree $T'$ such that $\mrca_{T'}(C_k) < \mrca_{T}(C_k)$.
\label{lemma:mrca_move}
\end{lemma}

\begin{proof}
Let $T$ and $R$ be trees as defined in the lemma.
Notice that the only $\rnni$ move on $T$ that could decrease the rank of $\mrca_T(C_k)$ are $\nni$ moves on the edge between $\mrca_T(C_k)$ and its child (if this edge has length one), or a rank move between $\mrca_T(C_k)$ and the node with rank one less.
Therefore, we will now distinguish these two cases:
Either there is an edge $(v,w)$ between nodes $v = \mrca_T(C_k)$ and $w$ with rank $rank_T(w) = rank_T(v) - 1$, or there is no such edge.

In the latter case the only $\rnni$ move that could decrease the rank of $v$ is a rank move swapping the ranks of $v$ and $w$.

Let us now assume that there is an edge $(v,w)$ in $T$.
Let $T_v$ be the subtree that is rooted in the child of $v$ that is not $w$ and let $T_{w_1}$ and $T_{w_2}$ be the subtrees rooted in the children of $w$.
We are now going to explain why only one of these two $\nni$ moves changes the rank of $\mrca_T(C_k)$.
It is important to notice that $C_k$ either is the disjoint union of two of the clusters $C_1, \ldots, C_{k-1}$ or the union of one of those clusters with a single taxon that is not included in any of those clusters.
In general we can say that $C_k = S_1 \dot\cup S_2$ is the union of two subsets of taxon set of $T$ such each of these subsets is contained in one of the subtrees $T_v, T_{w_1}$, or $T_{w_2}$.
In fact, $S_1$ and $S_2$ are contained in two different of these subtrees.
Now it is easy to see that just one of the $\nni$ moves on $(v,w)$ (illustrated in Figure~\ref{fig:mrca_move}) decreases the rank of $\mrca_T(C_k)$.

\begin{figure}[H]
\centering
\includegraphics[width=0.5\textwidth]{mrca_move}
\vspace{12pt}
\caption{Two $\nni$ moves that are possible on the edge $(v,w)$ of the tree $T'$ with subtrees $T_v, T_{w_1}, T_{w_2}$ as described in the proof of Lemma~\ref{lemma:mrca_move}.}
\label{fig:mrca_move}
\end{figure}
\end{proof}

Note that the uniqueness of the $\rnni$ move in Line~\ref{alg:findpath:line:move_set_down} of the algorithm $\findpath$ also implies that the algorithm is deterministic.
The running time of the algorithm is quadratic in the number of taxa.

Because the algorithm returns a path between pairs of trees in $\rnni$, the length of the path approximates the $\rnni$ distance from above.
A natural question then is how accurate this approximation is.
We have computationally shown that the algorithm $\findpath$ finds the correct distance for trees up to six taxa.
Specifically, we have implemented $\findpath$ \autocite{Collienne2019} and compared the results with the true $\rnni$ distance as explained earlier in this section.

\subsection{$\csort$}
\label{section:alg_csort}

In this section we introduce an algorithm to compute paths between \emph{caterpillar trees}, which are trees where each internal node is adjacent to at least one leaf.
As a result caterpillar trees have only one \emph{cherry}, which is a pair of taxa that share their parent.
A path between two caterpillar trees that only consists of caterpillar trees is called a \emph{caterpillar path}.
We can identify a caterpillar tree $T = [\{x_1, x_2\}, \{x_1, x_2, x_3\}, \ldots, \{x_1, \ldots, x_n\}]$ with the list of its taxa $[x_1, x_2, \ldots, x_n]$, assuming that $x_1 < x_2$ as natural numbers (recall that the set of taxa is $\{1, \ldots, n\}$).

The algorithm $\csort$ (Algorithm~\ref{alg:csort}) is a natural modification of the classical Bubble Sort algorithm \autocite{Knuth1997-pi}.
A path $p$ from $T$ to $R= [x_1, x_2, \ldots, x_n]$ is computed iteratively so that after $k$ steps the last $k$ taxa of $T$ and $R$ coincide.
Specifically, in step $k$ ($k = n, \ldots, 3$) the parent of taxon $x_{k}$ is moved up by $\nni$ moves to the position it has in $R$.
Notice that there might be more than just one such move per step, which means that $p$ can be extended by more than one tree in each step.

\begin{algorithm}[H]
\caption{$\csort$($T,R$)}
\label{alg:csort}
\begin{algorithmic}[1]
\STATE $[x_1, x_2, \ldots, x_n] := R$, $T' := T$, $p = [T']$
\FOR {$k=n, \ldots, 3$}
    \FOR {$i=1, \ldots, k$}
        \IF {$T'[i] = x_k$}
            \STATE Update $T'$: swap $T'[i]$ and $T'[i+1]$
            \STATE $p = p + T'$
        \ENDIF
    \ENDFOR
\ENDFOR
\RETURN $p$
\end{algorithmic}
\end{algorithm}

The running time of $\csort$ is quadratic in $n$.

The path $p$ returned by $\csort$ is a shortest caterpillar path because every tree modification on $p$ reduces the number of inversions of taxa in $T$ and $R$.
Indeed, no $\rnni$ move on caterpillar trees can reduce the number of taxa inversions by more than one, assuming that inversions with both taxa of a cherry are counted only once, that is, if both pairs of taxa $(x, z)$ and $(y, z)$ appear in opposite orders in $T$ and $R$ and $(x,y)$ is a cherry then they are counted as one inversion.
Hence any caterpillar path every move along which reduced the number of inversions is a shortest caterpillar path.

It is not obvious that the path between $T$ and $R$ returned by $\csort$ has the least possible length among all $\rnni$ paths (not only caterpillar paths), that is, the length of $d(T, R)$.
This fact will be established in Theorem~\ref{thm:caterpillar_convex}.

Our Python implementation of this algorithm can be accessed at \autocite{Collienne2019}.

\subsection{$\mdtree$}
\label{section:alg_mdtree}

The idea behind this algorithm is to efficiently return a tree as far away from a given tree as possible.
As we will see in Section~\ref{section:diameter}, $\mdtree$ achieves this goal exactly for caterpillar trees.

$\mdtree$ (Algorithm~\ref{alg:max_dist_tree}) works as follows.
Given an arbitrary tree $T$, order its taxa in a list $L = [l_1, \ldots, l_n]$ so that the ranks of their parents are non-decreasing with respect to this order.
Note that this list is not uniquely determined by the tree.
$\mdtree$ constructs an output tree $R$ as follows:
Initially, $R$ only consists of two taxa $l_1$ and $l_2$, the most recent common ancestor of which will eventually be of rank $n - 1$.
The remaining taxa are iteratively added to $R$ in the reversed order of $L$ starting from $l_n$.
At every iteration, the taxon is attached to a new internal node created on one of the existing branches in $R$ so that this new attachment node has rank one.
Note that $R$ is not uniquely determined due to the non-deterministic choice of the attachment edge.

\begin{algorithm}[H]
\caption{$\mdtree(T)$}
\label{alg:max_dist_tree}
\begin{algorithmic}[1]
\STATE Construct list $L=[l_1, \ldots, l_n]$ of all taxa ordered with respect to the ranks of their parents in $T$
\label{alg:mdtree_lineL}
\STATE Build tree $R$ with two taxa $l_1, l_2$ that are children of the root
\FOR {$i = 3, \ldots, n$}
\STATE Extend $R$ by attaching $l_i$ to a newly created node of rank $1$ on an arbitrarily chosen edge
\label{alg:mdtree_lineAddTaxon}
\ENDFOR
\RETURN $R$
\end{algorithmic}
\end{algorithm}

Note that the list computed in Line~\ref{alg:mdtree_lineL} of $\mdtree$, as well as the attachment edge of the new taxon in Line~\ref{alg:mdtree_lineAddTaxon} are not uniquely determined.
Therefore, this algorithm is non-deterministic, which is an important observation that we will use for finding the radius of $\rnni$ in Section~\ref{section:diameter}.

\section{Geometry of $\rnni$}
\label{section:geometry}

The study of shortest paths in the $\rnni$ graph in this section is motivated by our aim to understand the geometry of $\rnni$ as a metric space.
We will employ the algorithms developed in the previous section to aid our analysis of shortest paths.
Those algorithms will, for example, enable us to prove that every pair of caterpillar trees $T$ and $R$ is connected by a caterpillar path of length $d(T,R)$ (Theorem~\ref{thm:caterpillar_convex} in Section~\ref{section:caterpillar_convex}).
In other words, the set of caterpillar trees is \emph{convex} in $\rnni$.
In Section~\ref{section:diameter} we will investigate the diameter and radius of the $\rnni$ space.
Interestingly, the \emph{diameter} of $\rnni$, that is the maximum distance $\Delta(\rnni) = \max \{d(T, R) \mid T, R \in \rnni\}$ between any two trees in the graph, equals its \emph{radius} defined as $\min\limits_T \max\limits_R d(T,R)$.
Next, we will consider the so-called Split Theorem \autocite{Gavryushkin2018-ol}, which states that every shared split of taxa is maintained along every shortest path.
We will give a counterexample to the Split Theorem in $\rnni$ and consider a variant of the Weak Split Theorem, which states that every shared split of taxa is maintained along a shortest path.
We will call this statement the Clade Theorem, as it is slightly more natural to work with ranked trees in terms clades rather than splits (see Conjecture~\ref{conjecture:cluster_theorem}).

We start with the following auxiliary results.

\begin{lemma}
$\Delta(\rnni) \leq \frac{(n-1)(n-2)}{2}$ for $n \geq 3$.
\label{lemma:diameter_bound}
\end{lemma}

\begin{proof}
Since the diameter is bounded from above by the maximum length of a path computed by $\findpath$, it is enough to find this maximum.
Let us assume that $T$ and $R = [C_1, \ldots, C_{n-2}]$ are trees for which $\findpath$ computes a path of maximum length.
It follows that in each step of $\findpath$ the most recent common ancestor of the cluster $C_k$ considered in that step is the root.
Thus, there are $n-1-k$ $\rnni$ operations necessary to move $C_k$ to its correct position.
Hence, the maximum length of a path computed by $\findpath$ is bounded by $\sum\limits_{i = 1}^{n-2} (n-1-i) = \frac{(n-2)(n-1)}{2}$.
\end{proof}

The following lemma relates distances between trees on $n$ and $n+1$ taxa and is an important tool for inductive arguments.
We use the notion $\parent_T(x)$ to refer to the node adjacent to taxon $x$ in tree $T$.
$T{\big|}_n$ denotes the restriction of tree $T$ to the set of taxa $\{1, \ldots, n\}$.
In particular if $T$ is a tree on taxa $\{1, \ldots, n+1\}$ the tree $T{\big|}_n$ is obtained by deleting taxon $n+1$ and suppressing the thereby created node of degree two.

\begin{lemma}
Let $T$ and $R$ be two trees on taxa $\{1, \ldots, n+1\}$.
Then $d(T{\big|}_n, R{\big|}_n) \leq d(T,R) - \delta$, where $\delta = |\rank_T(\parent_T(n+1)) - \rank_R(\parent_R(n+1))|$.
\label{lemma:distance_delete_taxon}
\end{lemma}

\begin{proof}
First observe that the rank of the internal node $\parent_T(n+1)$ can only be changed by performing a rank move that involves $\parent_T(n+1)$ or an $\nni$ move on an edge adjacent to $\parent_T(n+1)$.
Second observe that any $\rnni$ move can change the rank of $\parent_T(n+1)$ by at most one.
Hence an $\rnni$ move on $T$ can decrease the rank of $\parent_T(n+1)$ by at most one.

Let $p$ be a shortest path from $T$ to $R$ and $p{\big|}_n$ the path resulting from deleting taxon $n+1$ from all trees on $p$.
Then $p{\big|}_n$ is a path from $T{\big|}_n$ to $R{\big|}_n$.
Recall that $\delta$ is the difference in ranks between the parents of taxon $n+1$ in $T$ and $R$.
Combining this with the observations above, we conclude that $|p{\big|}_n| \leq |p| - \delta$.
Since $d(T{\big|}_n,R{\big|}_n) \leq |p{\big|}_n|$ and $|p| = d(T,R)$, the desired inequality follows.
\end{proof}

\subsection{The set of caterpillar trees}
\label{section:caterpillar_convex}

In this section we restrict our attention to the set of caterpillar trees.
These trees are of particular interest because they are used to prove the computing the $\nni$ distance is an NP-hard problem.
As we will see below, shortest paths between caterpillar trees in $\rnni$ differ from those in the classical $\nni$ space.
We will show later in this section that $\rnni$ distances between caterpillar trees can be computed in polynomial time.
Throughout this section we use the list representation of trees described in Section~\ref{section:alg_csort}.

Recall \autocite{Gavryushkin2018-ol} that to compute a shortest path between two caterpillar trees in $\nni$ sometimes requires first building a clade and then move the clade around the tree (see Figure~\ref{fig:NNI_vs_RNNI}).
In $\rnni$ however, the necessity of additional rank moves invalidates this strategy.
This is due to the fact that $\nni$ moves $\rnni$ are only allowed on edges of length one.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{NNI_vs_RNNI}
\vspace{12pt}
\caption{Paths between caterpillar trees $T$ and $R$: Solid arrows indicate paths in $\rnni$, dashed arrow is a move only possible in $\nni$.
Bottom path is a shortest $\rnni$ path.}
\label{fig:NNI_vs_RNNI}
\end{figure}

These observations motivate the following general result.
Every pair of trees $T, R$ in $\rnni$ are connected by a caterpillar path of length $d(T,R)$, that is, the set of caterpillar trees is convex in $\rnni$ (Theorem~\ref{thm:caterpillar_convex}).
Note that this is not true in the $\nni$ space (Figure~\ref{fig:NNI_vs_RNNI}).

We need the following lemma.
Recall that $\frac{(n-1)(n-2)}{2}$ is the upper bound of the diameter of $\rnni$ by Lemma~\ref{lemma:diameter_bound}.

\begin{lemma}
Let $T$ and $R$ be two caterpillar trees.
Then $d(T,R) = \frac{(n-1)(n-2)}{2}$ if and only if $d_c(T,R) = \frac{(n-1)(n-2)}{2}$, where $d_c$ is the length of a shortest caterpillar path.
\label{lemma:caterpillar_dist=diameter}
\end{lemma}

\begin{proof}
First note that $d(T,R) = \frac{(n-1)(n-2)}{2}$ implies $d_c(T,R) = \frac{(n-1)(n-2)}{2}$.
Indeed, in each step $i=0, \ldots, n-3$ of $\csort$ at most $n-2-i$ pairs of taxa swap positions, as in the worst case the taxon considered in step $k$ is moved from the cherry of the tree up to the position $n-k$.
It follows that the maximum length of a path computed by $\csort$ is
\begin{equation}
\sum\limits_{i=0}^{n-3} n-2-i = \frac{(n-1)(n-2)}{2}.
\label{csort_max}
\end{equation}

For proving the other direction of the statement, assume that $d_c(T,R) = \frac{(n-1)(n-2)}{2}$.
We prove that $d(T,R) = \frac{(n-1)(n-2)}{2}$ by induction on the number of taxa $n$.
The induction basis for $n=3$ taxa is trivial as the $\rnni$ graph is a triangle on three caterpillar trees in that case.

For the induction step we assume without loss of generality that $T$ is the caterpillar tree $[1, \ldots, n+1]$ and $R$ is a caterpillar tree such that $d_c(T, R) = \frac{n(n-1)}{2}$.
Consider the path $p = \csort(T, R)$ of this length.
Note that in this case taxon $n+1$ has to be in the cherry of $R$, as otherwise $\csort$ would find a path between $T$ and $R$ of length strictly less than $\frac{n(n-1)}{2}$, which can be shown by counting the number of moves as in (\ref{csort_max}) above.

Let us now consider the path $p\big|_n$ from $T\big|_n$ to $R\big|_n$, which is obtained by restricting every tree on $p$ to taxa $1, \ldots, n$ and removing identical trees.
Observe that this path is shorter than $p$ by the number of moves that involve taxon $n+1$.
Since this taxon is adjacent to the root in $T$ and part of the cherry in $R$, $\csort$ requires $n-1$ such moves.
Hence, $d_c(T{\big|}_n, R{\big|}_n) = \frac{n(n-1)}{2} - (n-1)$ = $\frac{(n-1)(n-2)}{2} = d(T{\big|}_n,R{\big|}_n)$, by induction.
Lemma~\ref{lemma:distance_delete_taxon} implies that $d(T{\big|}_n, R{\big|}_n) \leq d(T,R) - |\rank_T(\parent_T(n+1)) - \rank_R(\parent_R(n+1))| = d(T,R) - (n-1)$.
So $\frac{(n-1)(n-2)}{2} \leq d(T,R) - (n-1)$, which implies that $d(T, R) \geq \frac{n(n-1)}{2}$.
Thus $d(T, R) = \frac{n(n-1)}{2}$.
\end{proof}

\begin{theorem}
The set of caterpillar trees is convex in $\rnni$.
\label{thm:caterpillar_convex}
\end{theorem}

\proof
Note that it is suffice to prove that $d_c(T, R) = d(T, R)$ for an arbitrary pair of caterpillar trees $T$ and $R$.
Lemma~\ref{lemma:caterpillar_dist=diameter} implies that if $T$ and $R$ are at the maximal possible distance, that is, $d_c(T, R) = \frac{(n-1)(n-2)}{2}$.
Denote the latter number by $D$.

Assume now that $d_c(T, R) = D - 1$.
If we apply $\csort(T, R)$ in this case, there must be a step in the algorithm where the taxon that is moved up is not part of the cherry (of $T'$) at the beginning of this step.
Let $x$ be the first taxon that has this property and is considered at step $k$.
Then there must be a taxon $y$ that is immediately preceding $x$ in the tree at the beginning of step $k$.
Since $k$ is the first such step, $y$ must be preceding $x$ in tree $T$ as well.
Consider a tree $\hat T$ that is obtained from $T$ by exchanging $x$ and $y$.
Note that the caterpillar path from $\hat T$ to $R$ computed by $\csort$ passes $T$ and coincides with the path $\csort(T, R)$ from there on.
Clearly, $d_c(\hat T, R) = D$.

Assuming $d_c(T, R) < D$, we can iterate the construction above to find a tree $\hat T$ such that $d_c(\hat T, R) = D$ and the caterpillar path from $\hat T$ to $R$ computed by $\csort$ passes $T$ and coincides with the path $\csort(T, R)$ from there on.
Lemma~\ref{lemma:caterpillar_dist=diameter} implies that $d(\hat T, R) = D$.

Assume that $d(T, R) < d_c(T, R)$ and let $p$ be a path from $T$ to $R$ of length $d(T, R)$.
Consider a path $q$ obtained by concatenating paths $\csort(\hat T, T)$ and $p$.
Note that $q$ is a path from $\hat T$ to $R$ which is shorter than $\csort(\hat T, R)$.
Indeed, the two paths coincide between $\hat T$ and $T$ and $q$ is shorter between $T$ and $R$.
Since $d_c(\hat T, R) = D$, the existence of $q$ is a contradiction to Lemma~\ref{lemma:caterpillar_dist=diameter}.
\endproof

\subsection{Diameter and radius}
\label{section:diameter}

\textcite[Theorem~7]{Gavryushkin2018-ol} gave an upper bound for the diameter of $\rnni$ space, which we recall is denoted by $\Delta(\rnni)$.
They showed that $\Delta(\rnni) = n^2 - 3n - 5/8$.
In this section we improve that result and calculate the exact diameter and radius of the $\rnni$ space.

\begin{theorem}
$\Delta(\rnni) = \frac{(n-1)(n-2)}{2}$.
\label{thm:diameter}
\end{theorem}

\proof
Since $d_c([1\ldots,n], [n-1,n,n-2,n-3, \ldots, 1]) = \frac{(n-1)(n-2)}{2}$, the claim follows directly from from Theorem~\ref{thm:caterpillar_convex} and Lemma~\ref{lemma:diameter_bound}.
\endproof

We show in the rest of this section that the radius of $\rnni$ space coincides with its diameter.
The main tool to establish this result is the algorithm $\mdtree$ from Section~\ref{section:alg_mdtree}.
Recall that $\mdtree$ is a non-deterministic algorithm, which implies that the output tree is not uniquely defined by the input to the algorithm.

\begin{lemma}
Let $T$ be a caterpillar tree.
Then $d(T, R) = \Delta(\rnni)$ for every tree $R$ such that $R = \mdtree(T)$.
\label{lemma:max_dist_caterpillar}
\end{lemma}

\begin{proof}
We prove the lemma by induction on the number of taxa $n$.
The induction basis for $n = 3$ is trivial.

Assume now that $T = [1, \ldots, n+1]$.
In this case, the parent of taxon $n+1$ has rank $n$ in $T$ and rank $1$ in $R$.
Lemma~\ref{lemma:distance_delete_taxon} then implies that
\begin{equation}
d(T\big|_n,R\big|_n) \leq d(T,R) - (n-1).
\label{eqn:max_dist_cat_proof}
\end{equation}
Note that $R\big|_n = \mdtree (T\big|_n)$ for an appropriate choice of attachment points in the execution of $\mdtree$.
By induction hypothesis and Theorem~\ref{thm:diameter}, $d(T\big|_n,R\big|_n) = \frac{(n-1)(n-2)}{2}$.
So inequality~(\ref{eqn:max_dist_cat_proof}) becomes $d(T,R) \geq \frac{(n-1)(n-2)}{2} + n - 1 = \frac{n(n-1)}{2}$, which with Theorem~\ref{thm:diameter} gives the desired equality.
\end{proof}

The non-determinism of $\mdtree$ can be exploited to investigate what trees are at the maximal possible from a given caterpillar tree.
In fact, those trees can have an arbitrary (non-ranked) topology!
This is because in each step of the algorithm the next taxon can be added on any edge incident to a leaf in the running tree.
Therefore, if the input tree is a caterpillar tree, $\mdtree$ can be used to find a tree of pre-defined topology that is at the maximal possible distance from the input tree.
Because the distance between trees is invariant under permutations of taxa labels, we conclude that for every tree $R$ on $n$ taxa there exists a caterpillar tree $T$ such that $d(T,R) = \Delta(\rnni) = \frac{(n-1)(n-2)}{2}$.
This proves the following Theorem~\ref{thm:radius}.

\begin{theorem}
The radius and diameter of the $\rnni$ space coincide and are equal to $\frac{(n-1)(n-2)}{2}$.
\label{thm:radius}
\end{theorem}

\subsection{Cluster Theorem}
\label{section:cluster_theorem}

In this section we will make a further step towards understanding the complexity of computing shortest paths in $\rnni$.

\begin{theorem}
There are trees $T,R$ in $\nni$ sharing a cluster which is not shared by any intermediate tree on any shortest path from $T$ to $R$.
\label{thm:split_nni}
\end{theorem}

\begin{proof}
See \autocite{Li1996-zw}.
\end{proof}

In the proof of Theorem~\ref{thm:split_nni} presented in \autocite{Li1996-zw} an example of two trees $T$ and $R$ that share clusters $C$ and $X \setminus C$ but have no shortest path that preserves these cluster is given.
A simplified illustration of this example is provided in Figure~\ref{fig:NNI_NP_proof}.
\autocite{Li1996-zw} prove that on a shortest path between $T$ and $R$ in $\nni$ at first $T$ is transformed to a tree $T'$ with $\frac{n}{2}$ cherries, each of which contains one taxon of each of the clusters $C$ and $X \setminus C$.
Then $T'$ is sorted to a tree $R'$ containing the same cherries as $T'$ such that that resolving all these cherries of $R'$ results in $R$.
Specifically, \autocite{Li1996-zw} prove that there is no shortest path between $T$ and $R$ in $\nni$ where $C$ and $X \setminus C$ are present in all trees on that path.

The reason that this example cannot be transferred to $\rnni$ is that there are lots of additional rank moves needed on a path between $T$ ans $R$ similar to the one described above.
As in the example depicted in Figure~\ref{fig:NNI_vs_RNNI}, adding rank moves to the shortest $\nni$ path from $T$ to $R$ results in a path that is longer than a shortest $\rnni$ path.
Sorting the two caterpillar subtrees of $T$ and $R$ separately results in a shorter path in $\rnni$.

\begin{figure}[H]
\centering
\includegraphics[width=.8\textwidth]{NNI_NP_proof}
\vspace{12pt}
\caption{Example of a shortest path between two trees $T$ and $R$ in $\nni$.
In \autocite{Li1996-zw} it is proven that there is a labelling that ensures that there is no shortest path in $\nni$ where the clusters (dots vs.\ squares) shared between $T$ and $R$ are preserved.
Instead, the depicted path where $\frac{n}{2}$ cherries are built and then sorted before being resolved is a shortest path}
\label{fig:NNI_NP_proof}
\end{figure}

Since the example of a shortest $\nni$ described above, which was used to prove Theorem~\ref{thm:split_nni} in \autocite{Li1996-zw}, does not work in $\rnni$, a contrary statement was conjectured in \autocite{Gavryushkin2018-ol}.
For understanding this conjecture it is important to know the connection between edges in a tree and so-called \emph{splits}, which are bipartitions of the taxon set of a tree:
By deleting an edge $e$ of a tree $T$ one receives two graphs with labels in $A$ and $B$, respectively, such that $\{A,B\}$ is a partition of the taxon set of $T$.
Note that each edge induces such a bipartition, often written as $A|B$, and that both edges incident to the root of a tree induce the same partition.
Since these partitions are commonly referred to as splits, the following conjecture was named \emph{Split Theorem} in \autocite{Gavryushkin2018-ol}.

\begin{conjecture}[Split Theorem]
For the $\rnni$ graph the following statement holds:
If a partition of leaves given by an edge is present in two trees $T$ and $R$ then the partition is presented in every tree on every shortest path between $T$ and $R$.
\label{conjecture:split_theorem}
\end{conjecture}

We found a simple counterexample to this conjecture, provided in Figure~\ref{fig:splitthm_counterexample}.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{splitthm_counterexample}
\vspace{12pt}
\caption{The split $123|45$ is present in $T$ and $R$, but the path at the top is a shortest path (computed by $\csort$) where none of the trees contains this split.
On the path at the bottom, which is a shortest path as well, this split is maintained.}
\label{fig:splitthm_counterexample}
\end{figure}

Since the counterexample in Figure~\ref{fig:splitthm_counterexample} shows that the version of the Split Theorem stated in \autocite{Gavryushkin2018-ol} does not hold, we will now claim an alternative to this conjecture, the \emph{Cluster Theorem} (Conjecture~\ref{conjecture:cluster_theorem}).
The Cluster Theorem concerns clusters instead of splits.
This is motivated by the fact that rooted phylogenetic trees can uniquely be represented by sets of clusters \autocite{Steel2016-ye}, but not by sets of splits as they cannot define the position of the root.
Note that trees $T$ and $R$ in Figure~\ref{fig:splitthm_counterexample} induce the same set of splits, but share no cluster.

\begin{conjecture}[Cluster Theorem]
For the $\rnni$ graph the following statement holds:
if two trees $T$ and $R$ contain the same cluster $C$, then $C$ is present as cluster in every tree on every shortest path between $T$ and $R$.
\label{conjecture:cluster_theorem}
\end{conjecture}

For providing further evidence that the Cluster Theorem holds in $\rnni$, we could show computationally that it holds for the $\rnni$ graph on small trees with up to seven taxa.
We did this by computing subgraph of $\rnni$ only containing trees that share a cluster and compare distances within this subgraph with distances in the whole $\rnni$ graph \autocite{Collienne2019}.
We computed these graphs according to the algorithm of \autocite{Gavryushkin2018-ol} as discussed in Section~\ref{section:algorithms}.

\section{Discussion}

Since the problem of computing distances in the most common tree spaces as $\nni$ \autocite{Dasgupta2000-xa}, rooted $\spr$ \autocite{Bordewich2005-nx}, or $\tbr$ \autocite{Allen2001-ky} is known to be $\np$-hard, it is desirable to find a tree space that is less complex.
So far it is not known whether computing distances in $\rnni$ is $\np$-hard or not.
The natural approach to solve this question is to compare $\rnni$ and $\nni$ graph to see whether the proof of $\np$-hardness of $\nni$ applies to $\rnni$ as well.
We will explain why this is not the case by considering an example of shortest paths in $\nni$ that builds the key of the proof of its $\np$-hardness.
After explaining why this example cannot be transferred to $\rnni$ we will introduce the so-called Split Theorem (Conjecture~\ref{conjecture:split_theorem}), which has been conjectured in \autocite{Gavryushkin2018-ol}.
However, we present a counterexample disproving the Split Theorem in its original version and claiming the Cluster Theorem (Conjecture~\ref{conjecture:cluster_theorem}) as alternative version.

The proof of $\np$-completeness of $\nni$ in \autocite{Dasgupta2000-xa} is based on the fact that Theorem~\ref{thm:split_nni} holds for $\nni$.
The aim of this paper was to gain better understanding of the $\rnni$ space of ranked phylogenetic trees.
Before investigating geometrical properties of $\rnni$ we introduced some algorithms in Section~\ref{section:algorithms}).
One of the most important algorithms we introduced here is $\findpath$, which can be used for approximating $\rnni$ distances.
We could in particular show that this algorithm computes exact distances for all pairs of trees on up to six taxa.
This algorithm might be a useful tool for future investigations of the $\rnni$ space.

We used the algorithms introduced in Section~\ref{section:algorithms} to prove some geometric properties of the $\rnni$ graph such as diameter and radius.
Moreover, we discovered that the set of caterpillar trees is convex in $\rnni$ (Theorem~\ref{thm:caterpillar_convex}).
With this result and the algorithm $\csort$ for computing shortest caterpillar paths (Section~\ref{section:alg_csort}) we can compute $\rnni$ distances between caterpillar trees in polynomial time.

This already suggests that there is a difference in complexity of computing distances in $\rnni$ and $\nni$, since in $\nni$ there is no subspace known for which computing distances only takes polynomial time.
In Section~\ref{section:cluster_theorem} we supported the claim that computing $\rnni$ distances is not as hard as computing $\nni$ distances further by discussing the proof of $\np$-completeness of computing distances in $\nni$ and showing that this proof does not work for $\rnni$.
Specifically, we conjectured the Cluster Theorem (Conjetcure~\ref{conjecture:cluster_theorem}), which does not hold in $\nni$.
However, we strongly suspect that it holds in $\rnni$ and could prove that it does hold for trees on up to seven taxa \autocite{Collienne2019}.

\subsection{Open Problems}

In Section~\ref{section:cluster_theorem} we claimed that the Cluster Theorem holds for $\rnni$.
We could show that the previous version of it, the Split Theorem as it was conjectured in \autocite{Gavryushkin2018-ol}, is not true.
If the Cluster Theorem actually is true still remains an open question.
Proving it will be a big step towards finding the complexity of computing $\rnni$ distances.

This question of whether computing $\rnni$ distances is $\np$-hard is another problem for future research.
Though we did not aim to answer this question in this paper, we made some progress towards solving it by establishing some geometric properties of the $\rnni$ graph.
Comparing these results with what is known for the plain $\nni$ graph makes us assume that the complexity of computing shortest path in $\rnni$ might differ from $\nni$.
We also suspect that the fact that computing distances between caterpillar trees can be done in polynomial time might be the key for further investigation of the complexity of computing $\rnni$ distances.

\printbibliography

\end{document}